---
title: "Assignment 1_Reproducible Research Course_Coursera"
author: "Ronaldo Braga"
date: "Wednesday, February 11, 2015"
output: html_document
---

This is the Assignment 1 from Reproducible Research Course from Coursera.

**Congigring the Environmemt**

```{r}
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')

library(data.table)
library(ggplot2)
library(plyr) 
library(xtable)
library(VIM)

```

**Loading the Data**


```{r}

represdata <- read.csv('activity.csv', header = TRUE, sep = ",",
                       colClasses=c("numeric", "character", "numeric"))

represdata$date <- as.Date(represdata$date, format = "%Y-%m-%d")
represdata$interval <- as.factor(represdata$interval)
# Checking the data using str() method:
str(rdata)

steps_per_day <- aggregate(steps ~ date, represdata, sum)
colnames(steps_per_day) <- c("date","steps")
head(steps_per_day)
```

**The Histogram**

```{r}
ggplot(steps_per_day, aes(x = steps)) + 
  geom_histogram(fill = "green", binwidth = 1000) + 
  labs(title="Histogram of Steps Taken per Day", 
       x = "Number of Steps per Day", y = "Number of times in a day(Count)") + theme_bw() 

steps_mean   <- mean(steps_per_day$steps, na.rm=TRUE)
steps_median <- median(steps_per_day$steps, na.rm=TRUE)

```

**What is the average daily activity pattern?**

```{r}

spi <- aggregate(represdata$steps, 
                                by = list(interval = represdata$interval),
                                FUN=mean, na.rm=TRUE)
spi$interval <- 
        as.integer(levels(spi$interval)[spi$interval])
colnames(spi) <- c("interval", "steps")

ggplot(spi, aes(x=interval, y=steps)) +   
        geom_line(color="green", size=2) +  
        labs(title="Average Daily Activity Pattern", x="Interval", y="Number of steps") +  
        theme_bw()


```

**Counting missing values**

Total number of missing values in steps can be calculated using is.na() 
```{r}
missing_vals <- sum(is.na(represdata$steps))
missing_vals
```

The total number of  missing values are 2304.


**Are there differences in activity patterns between weekdays and weekends?**

```{r}

day <- function(date){
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}

#creating day column
represdata$day <- as.factor(sapply(represdata$date, day))


#time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

par(mfrow = c(2, 1))
for (dayType in c("weekday", "weekend")) {
    stepType <- aggregate(steps ~ interval, data = represdata, subset = represdata$day == 
        dayType, FUN = mean)
    plot(stepType, type='l', main = dayType)
}



```
