---
title: "Assignment 1"
author: "Ronaldo Braga"
date: "Wednesday, February 14, 2015"
output: html_document
---

This is the Assignment 1 ref. Reproducible Research Course (Coursera). 

**Preparing the R Environment**

```{r}
library(knitr)
opts_chunk$set(echo = TRUE, results = 'hold')
library(data.table)
library(ggplot2)

```

__Loading the data__


```{r}
represdata <- read.csv('activity.csv', header = TRUE, sep = ",",
                  colClasses=c("numeric", "character", "numeric"))

represdata$date <- as.Date(represdata$date, format = "%Y-%m-%d")
represdata$interval <- as.factor(represdata$interval)

```

__Checking the data__

```{r, echo=FALSE}      }
str(represdata)
```

**What is mean total number of steps taken per day?**

```{r, echo=FALSE} 
steps_by_day <- aggregate(steps ~ date, represdata, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")
rmean <- mean(steps_by_day$steps)
rmedian <- median(steps_by_day$steps)

```

**What is the average daily activity pattern?**

```{r, echo=FALSE}      }

spi <- aggregate(represdata$steps, 
                                by = list(interval = represdata$interval),
                                FUN=mean, na.rm=TRUE)
#convert to integers
##this helps in plotting
spi$interval <- 
        as.integer(levels(spi$interval)[spi$interval])
colnames(spi) <- c("interval", "steps")

ggplot(spi, aes(x=interval, y=steps)) +   
        geom_line(color="green", size=2) +  
        labs(title="Average Daily Activity Pattern", x="Interval", y="Number of steps") +  
        theme_bw()


```

**Counting missing values**

Total number of missing values in steps can be calculated using is.na() 
```{r}
missing_vals <- sum(is.na(represdata$steps))
missing_vals
```

The total number of  missing values are 2304.


**Are there differences in activity patterns between weekdays and weekends?**

```{r}


```